Carrier phase and amplitude estimation for phase shift keying using pilots and data

### 2014-06-04 paper.tex
minor edits

Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the sample mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilot symbols $\abs{P} = 0, \tfrac{L}{32}, \tfrac{L}{8}, \tfrac{L}{2}, L$.  When $\abs{P} \neq 0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}_\pi^2$.  Otherwise, when $\abs{P}=0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$ as in~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.  The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least square estimator.  The solid lines are the estimator MSEs predicted by Theorem~\ref{thm:normality}.   %(and by Theorem~2 from~\cite{McKilliam_leastsqPSKnoncoICASSP_2012} when $\abs{P} = 0$).  
The prediction is made by dividing the asymptotic covariance matrix by $L$.  The theorem accurately predicts the behaviour of the phase estimator when SNR is sufficiently large.  As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$, and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.  Our asymptotic theory cannot be expected to model the behaviour of the estimator for fixed $L$ as SNR goes to zero.  Similar observations have been made in the literature~\cite{ViterbiViterbi_phase_est_1983,6365847}.  As the SNR increases the variance of the estimators converge to that of the estimator where all symbols are pilots, i.e. $\abs{P} = L$.  As expected, the figures indicate that the accuracy of the estimator increases with the number of pilot symbols.  One interesting observation is that, for low SNR, the slope of the MSE curve when no pilot symbols exist is steeper than the slope when pilot symbols do exist.  This indicates that pilot symbols are particularly important when the SNR is low.

>>>>

Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the sample mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilot symbols $\abs{P} = 0, \tfrac{L}{32}, \tfrac{L}{8}, \tfrac{L}{2}, L$.  When $\abs{P} \neq 0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}_\pi^2$.  Otherwise, when $\abs{P}=0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$ as in~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.  The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least square estimator.  The solid lines are the estimator MSEs predicted by Theorem~\ref{thm:normality}.   %(and by Theorem~2 from~\cite{McKilliam_leastsqPSKnoncoICASSP_2012} when $\abs{P} = 0$).  
The prediction is made by dividing the asymptotic covariance matrix by $L$.  The theorem accurately predicts the behaviour of the phase estimator when SNR is sufficiently large.  As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$, and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.  Our asymptotic theory asserts the normality of the phase estimator as $L\to\infty$ and therefore cannot be expected to model the behaviour of the estimator for fixed $L$ as SNR goes to zero.  Similar observations have been made in the literature~\cite{ViterbiViterbi_phase_est_1983,6365847}.  As the SNR increases the variance of the estimators converge to that of the estimator where all symbols are pilots, i.e. $\abs{P} = L$.  The phase estimator is more accurate for smaller constellation sizes, i.e. smaller $M$. The value of SNR at which the accuracy of the estimator departs from that when all symbols are pilots increases with $M$.  The point of departure is approximately \unit[0]{dB}, \unit[6]{dB}, and \unit[13]{dB} for BPSK, QPSK and $8$-PSK respectively.  %As expected, the figures indicate that the accuracy of the estimator increases with the number of pilot symbols.  One interesting observation is that, for low SNR, the slope of the MSE curve when no pilot symbols exist is steeper than the slope when pilot symbols do exist.  This indicates that pilot symbols are particularly important when the SNR is low.

### 2014-06-04 response2.tex
minor edits

\item \textbf{Comment:} ``we assume that the time-offset has been previously handled'' $\to$ perhaps some motivation for this assumption and the consequences with respect to the system model would be in place \\
\textbf{Response:}  The assumption of an accurate time offset estimator is common in this literature~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990} and is unlikely to come as a surprise to the intended audience.  There exist numerous estimators for the time offset in the literature.  We have added the sentence

@@

\item \textbf{Comment:} although different figures are shown for M=2,4,8 (phase estimation), the effect of the constellation size is not addressed in the text  \\
\textbf{Response:}  We have included the following sentences to Section~\ref{sec:simulations}:

``As expected, the figures indicate that the accuracy of the estimator increases with the number of pilot symbols.  One interesting observation is that, for low SNR, the slope of the MSE curve when no pilot symbols exist is steeper than the slope when pilot symbols do exist.  This indicates that pilot symbols are particularly important when the SNR is low.''

@@

``As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$, and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.  Our asymptotic theory cannot be expected to model the behaviour of the estimator for fixed $L$ as SNR goes to zero.  Similar observations have been made in the literature~\cite{ViterbiViterbi_phase_est_1983,6365847}.''

>>>>

\item \textbf{Comment:} ``we assume that the time-offset has been previously handled'' $\to$ perhaps some motivation for this assumption and the consequences with respect to the system model would be in place \\
\textbf{Response:}  The assumption of an accurate time offset estimator is common in this literature~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.% and is unlikely to come as a surprise to the intended audience.  
There exist numerous estimators for the time offset in the literature.  We have added the sentence

@@

\item \textbf{Comment:} although different figures are shown for M=2,4,8 (phase estimation), the effect of the constellation size is not addressed in the text  \\
\textbf{Response:}  The following text is included in Section~\ref{sec:simulations}:

``As the SNR increases the variance of the estimators converge to that of the estimator where all symbols are pilots, i.e. $\abs{P} = L$.  The phase estimator is more accurate for smaller constellation sizes, i.e. smaller $M$. The value of SNR at which the accuracy of the estimator departs from that when all symbols are pilots increases with $M$.  The point of departure is approximately \unit[0]{dB}, \unit[6]{dB}, and \unit[13]{dB} for BPSK, QPSK and $8$-PSK respectively.''

% ``As expected, the figures indicate that the accuracy of the estimator increases with the number of pilot symbols.  One interesting observation is that, for low SNR, the slope of the MSE curve when no pilot symbols exist is steeper than the slope when pilot symbols do exist.  This indicates that pilot symbols are particularly important when the SNR is low.''

@@

``As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$, and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.  Our asymptotic theory asserts the normality of the phase estimator as $L\to\infty$ and therefore cannot be expected to model the behaviour of the estimator for fixed $L$ as SNR goes to zero.  Similar observations have been made in the literature~\cite{ViterbiViterbi_phase_est_1983,6365847}''
